---
title: "Univariate Hurst Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r, warning=FALSE, message=FALSE}
# load libraries
library(easypackages)
libraries("readxl","ggplot2","MASS","psych","here","heplots","patchwork","knitr","ggseg","ggsegExtra","tidyverse")
source(here("code","cohens_d.R"))

# FDR threshold to use
fdr_thresh = 0.05

# define paths
codepath = here("code")
phenopath  = here("pheno")
datapath = here("data", "mrc_aims","Hurst_dwt")
resultpath = here("results")

# load HCP annotations
hcp_annot = read.delim(file.path(codepath, "GlasserHCP_annot.txt"))

# make a plot of Glasser HCP parcellation
p = ggseg(mapping=aes(fill = area), position = "stacked", atlas = glasser, show.legend = FALSE) +
  scale_fill_brain("glasser", package="ggsegExtra")
# ggsave(filename = here("figures","hcp_glasser_figure.pdf"))
p
```

Load in the phenotype data that has all the demographics and clinical variables.

```{r, warning=FALSE, message=FALSE}
# Read in pheno file
pheno_file = file.path(codepath, "data","tidy_pheno_data.xlsx")
pheno_data = read_excel(pheno_file)
# change character columns to factors
pheno_data$Group = factor(pheno_data$Group)
pheno_data$Sex = factor(pheno_data$Sex)
pheno_data$Diagnosis = factor(pheno_data$Diagnosis)
```

Compute the descriptive statistics and camouflaging score

```{r, warning=FALSE, message=FALSE}
mask = pheno_data$use_subs==1
pheno_data_subset = subset(pheno_data, mask)
pheno_data_subset = data.frame(pheno_data_subset)
rownames(pheno_data_subset) = pheno_data_subset$subid

asd_data = subset(pheno_data_subset, pheno_data_subset$Diagnosis=="Autism")

# Standardize AQ
mean_AQ_ASC = mean(asd_data$AQ)
max_AQ = 50
asd_data$zAQ = (asd_data$AQ - mean_AQ_ASC)/max_AQ

# Standardize ADOS
mean_ADOS_SC = mean(asd_data$ADOS_CommSoc, na.rm = TRUE)
max_ADOS_SC = 22
asd_data$zADOS_SC = (asd_data$ADOS_CommSoc - mean_ADOS_SC)/max_ADOS_SC

# Standardize RMET
mean_RMET_ASC = mean(asd_data$RMET)
max_RMET = 36
asd_data$zRMET = (asd_data$RMET - mean_RMET_ASC)/max_RMET

# compute difference in standardized scores
asd_data$CF1 = asd_data$zAQ - asd_data$zADOS_SC
asd_data$CF2 = (asd_data$zRMET*-1) - asd_data$zADOS_SC

# run pca on CF1 and CF2
data4pca = asd_data[!is.na(asd_data$CF1),c("sub_id","CF1","CF2")]
pca_res = princomp(~CF1+CF2, data = data4pca, na.action = na.omit)

# insert CF_PC1 into data
data4pca$CF_PC1 = pca_res$scores[,1]
asd_data = merge(asd_data, data4pca[,c("sub_id","CF_PC1")], by = "sub_id")

# summarize PCA on CF1 and CF2
summary(pca_res)

# ANOVA for difference between autistic females vs males on CF_PC1
mod2use = lm(CF_PC1 ~ Sex, data = asd_data)
anova(mod2use)

# effect size difference in CF_PC1
es_res = cohens_d(asd_data[asd_data$Sex=="F","CF_PC1"],asd_data[asd_data$Sex=="M","CF_PC1"])
es_res

# merge asd_data back into the pheno_data_subset
pheno_data_subset$CF1 = NA
pheno_data_subset$CF2 = NA
pheno_data_subset$CF_PC1 = NA
sublist = asd_data$sub_id
for (isub in 1:length(sublist)){
  mask = pheno_data_subset$sub_id==asd_data$sub_id[isub]
  pheno_data_subset$CF1[mask] = asd_data$CF1[isub]
  pheno_data_subset$CF2[mask] = asd_data$CF2[isub]
  pheno_data_subset$CF_PC1[mask] = asd_data$CF_PC1[isub]
}

# variables to describe
vars2use = c("Diagnosis","Sex","Age","VIQ","PIQ","FIQ","CF_PC1","AQ","ADI_Social","ADI_Comm","ADI_Rep","ADOS_Comm","ADOS_Social","ADOS_Rep","ADOS_CommSoc","RMET","meanFD")

# get descriptive statistics by Diagnosis
describeBy(x = pheno_data_subset[,vars2use], group = c("Diagnosis","Sex"))
```

Analyze meanFD, age, and IQ. 

```{r, warning=FALSE, message=FALSE}
# Analysis on meanFD 
mod2use = lm(meanFD ~ Diagnosis*Sex, data = pheno_data_subset)
anova(mod2use)

# Analysis of Age
mod2use = lm(Age ~ Diagnosis*Sex, data = pheno_data_subset)
anova(mod2use)

# Analysis of VIQ
mod2use = lm(VIQ ~ Diagnosis*Sex, data = pheno_data_subset)
anova(mod2use)

# Subset only TD subjects and look for sex difference
pheno_data_subset_td = subset(pheno_data_subset, pheno_data_subset$Diagnosis=="TD")
mod2use = lm(VIQ ~ Sex, data = pheno_data_subset_td)
anova(mod2use)

# Subset only ASD subjects and look for sex difference
pheno_data_subset_asd = subset(pheno_data_subset, pheno_data_subset$Diagnosis=="Autism")
mod2use = lm(VIQ ~ Sex, data = pheno_data_subset_asd)
anova(mod2use)

# Analysis of PIQ
mod2use = lm(PIQ ~ Diagnosis*Sex, data = pheno_data_subset)
anova(mod2use)

# Analysis of FIQ
mod2use = lm(FIQ ~ Diagnosis*Sex, data = pheno_data_subset)
anova(mod2use)

# Analysis of RMET within Autism
asd_pheno_subset = subset(pheno_data_subset, pheno_data_subset$Diagnosis=="Autism")
mod2use = lm(RMET ~ Sex, data = asd_pheno_subset)
anova(mod2use)

# Analysis of ADI_Social
mod2use = lm(ADI_Social ~ Sex, data = asd_pheno_subset)
anova(mod2use)
```

# Analyze Hurst exponent data

```{r, warning=FALSE, message=FALSE}
ncols2use = dim(hcp_annot)[1]

# make temporary parcel labels
region_labels = character(length=0)
for (i in 1:dim(hcp_annot)[1]) {
  region_labels = c(region_labels, sprintf("parcel_%03d",i))
}# for i

# load Hurst data
Hdata = read.csv(file.path(codepath, "data", "tidy_H_data.csv"),row.names=1)

# merge Hurst data to pheno data
pheno_data_subset = merge(pheno_data_subset, Hdata, by = "sub_id")

# make data frames for each autism group to use later in clinical correlation analysis
asd_data = subset(pheno_data_subset,pheno_data_subset$Diagnosis=="Autism")
asdf_data = subset(asd_data,asd_data$Sex=="F")
asdm_data = subset(asd_data,asd_data$Sex=="M")

td_data = subset(pheno_data_subset,pheno_data_subset$Diagnosis=="TD")
tdf_data = subset(td_data,td_data$Sex=="F")
tdm_data = subset(td_data,td_data$Sex=="M")

# initialize output_res data frame to store results into
colnames2use = c("Sex_Fstat", "Sex_pval","Sex_fdrq","Sex_EffectSize_MvsF",
                 "Dx_Fstat", "Dx_pval","Dx_fdrq","Dx_EffectSize_TDvsASD",
                 "Dx_x_Sex_Fstat", "Dx_x_Sex_pval","Dx_x_Sex_fdrq","Dx_x_Sex_partialEtaSq",
                 "r_ASDf_CF","pval_ASDf_CF","fdrq_ASDf_CF",
                 "r_ASDm_CF","pval_ASDm_CF","fdrq_ASDm_CF",
                 "z_ASDf_vs_ASDm_CF","pval_ASDf_vs_ASDm_CF")
output_res = data.frame(matrix(nrow = ncols2use, ncol = length(colnames2use)))
colnames(output_res) = colnames2use
rownames(output_res) = region_labels

# test each parcel for between-group difference in Diagnosis
for (icol in 1:ncols2use) {
  # make formula
  form2use = as.formula(sprintf("%s ~ Diagnosis*Sex + meanFD + FIQ", region_labels[icol]))

  # run linear model
  mod2use = lm(formula = form2use, data = pheno_data_subset)
  
  # run an ANOVA
  res2use = anova(mod2use)
  
  # grab ANOVA results and store into output_res
  output_res[icol, "Sex_Fstat"] = res2use["Sex", "F value"]
  output_res[icol, "Sex_pval"] = res2use["Sex", "Pr(>F)"]
  output_res[icol, "Dx_Fstat"] = res2use["Diagnosis", "F value"]
  output_res[icol, "Dx_pval"] = res2use["Diagnosis", "Pr(>F)"]
  output_res[icol, "Dx_x_Sex_Fstat"] = res2use["Diagnosis:Sex", "F value"]
  output_res[icol, "Dx_x_Sex_pval"] = res2use["Diagnosis:Sex", "Pr(>F)"]
  
  # compute effect size for diagnosis main effect
  es_res = cohens_d(pheno_data_subset[pheno_data_subset$Diagnosis=="TD",region_labels[icol]], pheno_data_subset[pheno_data_subset$Diagnosis=="Autism",region_labels[icol]])
  output_res[icol, "Dx_EffectSize_TDvsASD"] = es_res
  
  # compute effect size for sex main effect
  es_res = cohens_d(pheno_data_subset[pheno_data_subset$Sex=="M",region_labels[icol]], pheno_data_subset[pheno_data_subset$Sex=="F",region_labels[icol]])
  output_res[icol, "Sex_EffectSize_MvsF"] = es_res
  
  # compute partial eta-squared as the effect size of the interaction effect
  eta_sq_res = etasq(mod2use)
  output_res[icol, "Dx_x_Sex_partialEtaSq"] = eta_sq_res["Diagnosis:Sex","Partial eta^2"]
  
  # analyze clinical correlations
  
  # correlations within autism males
  tmp_res = cor.test(asdm_data$CF_PC1,asdm_data[,region_labels[icol]])
  output_res[icol, "r_ASDm_CF"] = tmp_res$estimate
  output_res[icol, "pval_ASDm_CF"] = tmp_res$p.value

  # correlations within autism females
  tmp_res = cor.test(asdf_data$CF_PC1,asdf_data[,region_labels[icol]])
  output_res[icol, "r_ASDf_CF"] = tmp_res$estimate
  output_res[icol, "pval_ASDf_CF"] = tmp_res$p.value

  # test if the correlations in autism males and females is different
  r_diff = paired.r(xy = output_res[icol,"r_ASDf_CF"], 
           xz = output_res[icol,"r_ASDm_CF"],
           n = dim(asdf_data)[1],
           n2 = dim(asdm_data)[1])
  output_res[icol, "z_ASDf_vs_ASDm_CF"] = r_diff$z
  output_res[icol, "pval_ASDf_vs_ASDm_CF"] = r_diff$p
  
}# for icol
output_res = output_res[-c(131,172),]
hcp_annot = hcp_annot[-c(131,172),]

# compute FDR for all effects
output_res$Dx_fdrq = p.adjust(output_res$Dx_pval, method = "fdr")
output_res$Sex_fdrq = p.adjust(output_res$Sex_pval, method = "fdr")
output_res$Dx_x_Sex_fdrq = p.adjust(output_res$Dx_x_Sex_pval, method = "fdr")
output_res$fdrq_ASDm_CF = p.adjust(output_res$pval_ASDm_CF, method = "fdr")
output_res$fdrq_ASDf_CF = p.adjust(output_res$pval_ASDf_CF, method = "fdr")

# make columns for parcel_index and area
output_res$parcel_index = rownames(output_res)
output_res$area = hcp_annot$area
output_res$area_descrip = hcp_annot$area_descrip

# redo rownames of the output_res to be the area names
rownames(output_res) = hcp_annot$area
# write.csv(output_res,file = file.path(resultpath,"Hurst_output_res.csv"))
```

Show sex*diagnosis interaction

```{r, warning=FALSE, message=FALSE}
tmp_res = output_res[order(output_res$Dx_x_Sex_fdrq),]
vars2use = c("area_descrip","Dx_x_Sex_Fstat","Dx_x_Sex_pval","Dx_x_Sex_fdrq","Dx_x_Sex_partialEtaSq")
kable(tmp_res[,vars2use])
```

Plot sex*diagnosis interaction effect.

```{r, warning=FALSE, message=FALSE}
yax_label = "Hurst exponent (H)"

# show effect size throughout the entire brain
g1 = ggseg(.data=output_res, atlas=glasser, position="stacked",mapping=aes(fill=Dx_x_Sex_partialEtaSq)) +
  theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Partial eta-sq"))+
  scale_fill_gradient2(midpoint=0.13/2, low="blue", mid="white", high="red", limits=c(0, 0.13), breaks = round(seq(from=0,to=0.12,by=0.03),2)) + guides(fill = guide_colourbar(nbin = 100))

# g1 = ggseg(data = output_res, mapping=aes(fill = Dx_x_Sex_partialEtaSq), position = "stacked", atlas = "glasser") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Partial eta-sq")) +
#   scale_fill_gradient2(midpoint=0.13/2, low="blue", mid="white", high="red", limits=c(0, 0.13), breaks = round(seq(from=0,to=0.12,by=0.03),2)) + guides(fill = guide_colourbar(nbin = 100))
# ggsave(filename = here("figures","sex_x_dx_interaction_Hurst_wholebrain.pdf"), plot = g1)
g1

# show the regions passing FDR q<0.05
output_res_subset = subset(output_res, output_res$Dx_x_Sex_fdrq<=0.05)
g2 = ggseg(.data=output_res_subset, atlas=glasser, position="stacked",mapping=aes(fill=Dx_x_Sex_partialEtaSq)) +
  theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Partial eta-sq"))+
  scale_fill_gradient2(midpoint=0.13/2, low="blue", mid="white", high="red", limits=c(0, 0.13), breaks = round(seq(from=0,to=0.12,by=0.03),2)) + guides(fill = guide_colourbar(nbin = 100))

# g2 = ggseg(data = output_res_subset, mapping=aes(fill = Dx_x_Sex_partialEtaSq), position = "stacked", atlas = "glasser") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Partial eta-sq")) +
#   scale_fill_gradient2(midpoint=0.13/2, low="blue", mid="white", high="red", limits=c(0, 0.13), breaks = round(seq(from=0,to=0.12,by=0.03),2)) + guides(fill = guide_colourbar(nbin = 100))
# ggsave(filename = here("figures","sex_x_dx_interaction_Hurst_fdr.pdf"), plot = g2)
g2

# Region p32
area_name = "p32"
area_descrip = "Area p32"
reg_name = output_res$parcel_index[output_res$area==area_name]

cols2use = wesanderson::wes_palette("Moonrise2")[c(1,3)] 

# scatter-boxplots
p1 = ggplot(data = pheno_data_subset, aes_string(x = "Diagnosis", y = reg_name, colour = "Diagnosis")) + facet_grid(. ~ Sex)
p1 = p1 + geom_jitter() + geom_boxplot(fill = NA, colour = "#000000", outlier.shape = NA) + 
  guides(colour = FALSE) + ylab(yax_label) + scale_colour_manual(values=cols2use) +
  ggtitle(area_descrip) + theme(plot.title = element_text(hjust = 0.5))
# ggsave(filename = here("figures","sex_x_dx_interaction_Hurst_p32.pdf"), plot = p1)
p1
```

Describe results for Area p32

```{r, warning=FALSE, message=FALSE}
area_name = "p32"
area_descrip = "Area p32"
reg_name = output_res$parcel_index[output_res$area==area_name]

# male TD vs ASD difference
male_subset = subset(pheno_data_subset, pheno_data_subset$Sex=="M")
male_dx_effect_size = cohens_d(male_subset[male_subset$Diagnosis=="TD",reg_name],male_subset[male_subset$Diagnosis=="Autism",reg_name])
male_dx_effect_size

# female TD vs ASD difference
female_subset = subset(pheno_data_subset, pheno_data_subset$Sex=="F")
female_dx_effect_size = cohens_d(female_subset[male_subset$Diagnosis=="TD",reg_name],female_subset[male_subset$Diagnosis=="Autism",reg_name])
female_dx_effect_size
```

# Correlations with camouflaging

```{r, warning=FALSE, message=FALSE}
region_names = c("p32")
mask = is.element(output_res$area,region_names)
vars2use = c("area_descrip","r_ASDf_CF","pval_ASDf_CF","fdrq_ASDf_CF",
             "r_ASDm_CF","pval_ASDm_CF","fdrq_ASDm_CF",
             "z_ASDf_vs_ASDm_CF","pval_ASDf_vs_ASDm_CF")
kable(output_res[mask,vars2use])

cols2use = c(wesanderson::wes_palette("Zissou1")[c(1)],wesanderson::wes_palette("Darjeeling1")[c(4)]) 
xax_label = "Camouflaging Score"

# Region p32
area_name = "p32"
area_descrip = "Area p32"
reg_name = output_res$parcel_index[output_res$area==area_name]

# ggseg plot to show the brain region
output_res_subset = subset(output_res, output_res$area==area_name)
g1 = ggseg(.data = output_res_subset, mapping=aes(fill = r_ASDf_CF), 
           position = "stacked", atlas = glasser, show.legend = FALSE) + 
  ggtitle(area_descrip) + theme(plot.title = element_text(hjust = 0.5))
# g1 = ggseg(data = output_res_subset, mapping=aes(fill = r_ASDf_CF), 
#            position = "stacked", atlas = "glasser", show.legend = FALSE) + 
#   ggtitle(area_descrip) + theme(plot.title = element_text(hjust = 0.5))

# Scatterplot
p1 = ggplot(data = asd_data, aes_string(x = "CF_PC1", y = reg_name, colour = "Sex", fill = "Sex"))
p1 = p1 + geom_point() + geom_smooth(method = lm) + ylab(yax_label) + xlab(xax_label) + scale_color_manual(values=cols2use) + scale_fill_manual(values=cols2use) + guides(colour = FALSE, fill=FALSE) +  
  ggtitle(area_descrip) + theme(plot.title = element_text(hjust = 0.5))

p_final = g1 + p1 + plot_layout(nrow = 1, ncol = 2) 
# ggsave(filename = here("figures","camouflaging_correlations_Hurst_p32.pdf"), plot = p_final)
p_final
```


# Sex differences

```{r, warning=FALSE, message=FALSE}
tmp_res = output_res[order(output_res$Sex_fdrq),]
vars2use = c("area_descrip","Sex_Fstat","Sex_pval","Sex_fdrq","Sex_EffectSize_MvsF")
kable(tmp_res[,vars2use])

p1 = ggseg(.data = output_res, mapping=aes(fill=Sex_EffectSize_MvsF), position="stacked", atlas=glasser) + 
  ggtitle("Sex Difference (unthresholded)") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Cohen's d")) +
  scale_fill_gradient2(midpoint=0.35, low="blue", mid="white", high="red", limits=c(0, 1), breaks = round(seq(from=0,to=1,by=0.2),1)) + guides(fill = guide_colourbar(nbin = 100))
# p1 = ggseg(data = output_res, mapping=aes(fill=Sex_EffectSize_MvsF), position="stacked", atlas="glasser") + 
#   ggtitle("Sex Difference (unthresholded)") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Cohen's d")) +
#   scale_fill_gradient2(midpoint=0.35, low="blue", mid="white", high="red", limits=c(0, 0.85), breaks = round(seq(from=0,to=0.8,by=0.2),1)) + guides(fill = guide_colourbar(nbin = 100))

# Sex main effect regions surviving FDR q<0.05
output_res_subset = subset(output_res, output_res$Sex_fdrq<=0.05)
p2 = ggseg(.data = output_res_subset, 
           mapping=aes(fill=Sex_EffectSize_MvsF), 
           position="stacked", 
           atlas=glasser) + 
  ggtitle("Sex Difference (FDR q<0.05)") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill=guide_legend(title="Cohen's d")) +
  scale_fill_gradient2(midpoint=0.35, low="blue", mid="white", high="red", limits=c(0, 1), breaks = round(seq(from=0,to=1,by=0.2),1)) + 
  guides(fill = guide_colourbar(nbin = 100))
# p2 = ggseg(data = output_res_subset, mapping=aes(fill=Sex_EffectSize_MvsF), position="stacked", atlas="glasser") + 
#   ggtitle("Sex Difference (FDR q<0.05)") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Cohen's d")) +
#   scale_fill_gradient2(midpoint=0.35, low="blue", mid="white", high="red", limits=c(0, 0.85), breaks = round(seq(from=0,to=0.8,by=0.2),1)) + guides(fill = guide_colourbar(nbin = 100))

p_final = p1 + p2 + plot_layout(nrow=2, ncol=1)
# ggsave(filename = file.path(here(),"figures","sex_maineffects_unthresholded_Hurst.pdf"), plot = p_final)
p_final
```

# Diagnosis main effects

```{r, warning=FALSE, message=FALSE}
tmp_res = output_res[order(output_res$Dx_fdrq),]
vars2use = c("area_descrip","Dx_Fstat","Dx_pval","Dx_fdrq","Dx_EffectSize_TDvsASD")
kable(tmp_res[,vars2use])

p = ggseg(.data = output_res, 
          mapping=aes(fill=Dx_EffectSize_TDvsASD), 
          position="stacked", 
          atlas=glasser) + 
  ggtitle("Diagnosis Main Effect (unthresholded)") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill=guide_legend(title="Cohen's d")) + 
  scale_fill_gradient2(midpoint=0, low="blue", mid="white", high="red", limits=c(-0.65, 0.65), breaks = round(seq(from=-0.6,to=0.6,by=0.3),1)) + 
  guides(fill = guide_colourbar(nbin = 100)) 
# p = ggseg(data = output_res, mapping=aes(fill=Dx_EffectSize_TDvsASD), position="stacked", atlas="glasser") + 
#   ggtitle("Diagnosis Main Effect (unthresholded)") + theme(plot.title = element_text(hjust = 0.5)) + guides(fill=guide_legend(title="Cohen's d")) + scale_fill_gradient2(midpoint=0, low="blue", mid="white", high="red", limits=c(-0.65, 0.65), breaks = round(seq(from=-0.6,to=0.6,by=0.3),1)) + guides(fill = guide_colourbar(nbin = 100)) 

# ggsave(filename = file.path(here(),"figures","diagnosis_maineffects_unthresholded_Hurst.pdf"), plot = p)
p
```
