---
title: "Overlap between ASD Genes and DHT DE genes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
easypackages::libraries("here","ggplot2","gplots")
code_dir = here("code")
source(file.path(code_dir,"genelistOverlap.R"))

genelist_dir = here("EI_hurst","data","asd_genelists")

backgroundTotal = 13284
dht_de_genes = read.delim(file.path(genelist_dir,"dysreg_sv_limma_FDR_q0.05_resultTable.txt"))
dht_de_genes = as.character(dht_de_genes$GeneSymbol)

# ASD risk genes from Satterstrom et al., 2020, Cell
satterstrom_102_Ex = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_ExGenes.txt"),header = FALSE)$V1)
satterstrom_102_In = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_InGenes.txt"),header = FALSE)$V1)
satterstrom_102_MG = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_MicrogliaGenes.txt"),header = FALSE)$V1)
satterstrom_102_AsOPC = as.character(read.csv(file.path(genelist_dir,"satterstrom_2020_cell_102asdriskgenes_AsOPCGenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in excitatory neurons
excitatory_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_L23456degenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in inhibitory neurons
inhibitory_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_INPVSSTVIPSV2Cdegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in Microglia
microglia_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_Microgliadegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in Oligodendrocytes
oligodendrocyte_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_Oligodendrocytesdegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in Astrocytes
astrocyte_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_Astrocytedegenes.txt"),header = FALSE)$V1)

# genes from Velmeshev et al., 2019, Science, single nucleus RNA-seq DE genes in OPCs
opc_de_genes = as.character(read.csv(file.path(genelist_dir,"velmeshev_2019_science_OPCdegenes.txt"),header = FALSE)$V1)
```

## Overlap between Excitatory neurons in Velmeshev et al., 2019, Science AND any excitatory gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
ExRisk_de_genes = unique(c(excitatory_de_genes,
                           satterstrom_102_Ex))
overlap_ExRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(ExRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_ExRisk_res[[1]]$OR
overlap_ExRisk_res[[1]]$hypergeo_p
sort(overlap_ExRisk_res[[1]]$overlapping_genes)
```

## Overlap between Inhibitory neurons in Velmeshev et al., 2019, Science AND any Inhibitory gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
InRisk_de_genes = unique(c(inhibitory_de_genes,
                           satterstrom_102_In))
overlap_InRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(InRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_InRisk_res[[1]]$OR
overlap_InRisk_res[[1]]$hypergeo_p
sort(overlap_InRisk_res[[1]]$overlapping_genes)
```

## Overlap between Microglia cells in Velmeshev et al., 2019, Science AND any Microglia gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
MGRisk_de_genes = unique(c(microglia_de_genes,
                           satterstrom_102_MG))
overlap_MGRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(MGRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_MGRisk_res[[1]]$OR
overlap_MGRisk_res[[1]]$hypergeo_p
sort(overlap_MGRisk_res[[1]]$overlapping_genes)
```

## Overlap between Astrocyte, OPC, or Oligodendrocyte cells in Velmeshev et al., 2019, Science AND any Astrocyte and OPC gene from Satterstrom

```{r, warning=FALSE, message=FALSE}
AsOPCRisk_de_genes = unique(c(oligodendrocyte_de_genes,
                              opc_de_genes,
                              astrocyte_de_genes,
                              satterstrom_102_AsOPC))
overlap_AsOPCRisk_res = genelistOverlap(data.frame(dht_de_genes),
                                data.frame(AsOPCRisk_de_genes),
                                backgroundTotal,
                                print_result = FALSE,
                                header = FALSE)
overlap_AsOPCRisk_res[[1]]$OR
overlap_AsOPCRisk_res[[1]]$hypergeo_p
sort(overlap_AsOPCRisk_res[[1]]$overlapping_genes)
```

## Make a heatmap of overlapping DHT and autism-associated excitatory genes

```{r, warning=FALSE, message=FALSE}
df = read.csv(here("EI_hurst","data","overlap_dht_de_data.csv"),row.names=1)
df = as.matrix(df)

# plot the heatmap of the data matrix
heatmap.2(df,
          Rowv=TRUE, Colv=TRUE, 
          col=bluered,
          scale="row", 
          trace="none",
          srtCol = 25)
```